On the properties of projectile motion and quadratic air resistance

### 2016-02-20 Paper/main.tex
Spelling and new ref

	The time to highest point is the amount of time it takes for our object to reach it's peak height which we will denote $t_{peak}$. Since $y(t)$ is a second degree polynominal it only has one extremum. This must be the maximum because $v_y(0) > 0$. We can therefore solve $v_y = 0$ for $t_{peak}$.

@@

	We can plot another $y(t)$ where $k$ is much smaller which should result in our object not reaching it's terminal velocity in such short time. The following measurements are from a small steel ball with $k = 0.00313$.
	\begin{align}\nonumber
		\begin{cases}\nonumber
			C_D &= 0.47 \\
			m &= 0.022\,kg\\
			y_0 &= 3\,m\\
			v_0 &= 0\,m/s\\
			\rho &= 1.225\,kg/m^3\\
			A &= 0.00023916\,m^2
		\end{cases}
	\end{align}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ xlabel=t (s), ylabel=y (m)]
		\addplot[solid] coordinates {(0.000, 3.000) (0.010, 3.000) (0.020, 2.998) (0.030, 2.996) (0.040, 2.992) (0.050, 2.988) (0.060, 2.982) (0.070, 2.976) (0.080, 2.969) (0.090, 2.960) (0.100, 2.951) (0.110, 2.941) (0.120, 2.929) (0.130, 2.917) (0.140, 2.904) (0.150, 2.890) (0.160, 2.874) (0.170, 2.858) (0.180, 2.841) (0.190, 2.823) (0.200, 2.804) (0.210, 2.784) (0.220, 2.762) (0.230, 2.740) (0.240, 2.717) (0.250, 2.693) (0.260, 2.668) (0.270, 2.642) (0.280, 2.615) (0.290, 2.587) (0.300, 2.558) (0.310, 2.528) (0.320, 2.497) (0.330, 2.466) (0.340, 2.433) (0.350, 2.399) (0.360, 2.364) (0.370, 2.328) (0.380, 2.292) (0.390, 2.254) (0.400, 2.215) (0.410, 2.175) (0.420, 2.135) (0.430, 2.093) (0.440, 2.050) (0.450, 2.007) (0.460, 1.962) (0.470, 1.917) (0.480, 1.870) (0.490, 1.823) (0.500, 1.774) (0.510, 1.725) (0.520, 1.674) (0.530, 1.623) (0.540, 1.570) (0.550, 1.517) (0.560, 1.463) (0.570, 1.407) (0.580, 1.351) (0.590, 1.294) (0.600, 1.236) (0.610, 1.176) (0.620, 1.116) (0.630, 1.055) (0.640, 0.993) (0.650, 0.930) (0.660, 0.866) (0.670, 0.801) (0.680, 0.735) (0.690, 0.668) (0.700, 0.600) (0.710, 0.531) (0.720, 0.461) (0.730, 0.391) (0.740, 0.319) (0.750, 0.246) (0.760, 0.172) (0.770, 0.098) (0.780, 0.022) (0.790, -0.055) };

@@

	In contrary to our LAT approximation in section \ref{seq:lat} we must now separate the cases where $\hat{w} = \hat{\jmath}, \hat{w} = -\hat{\jmath}$ because a negative $w$ should not lead to a positive $u'$. When $v$ was approximated as $u$ we assumed that $\hat{u} = \hat{\imath}$ because it is the only realistic scenario. Now that $v$ is approximated as $w$ we need to consider it's direction before that context. In the end our equations of motions will consist of two separate $u$ and $w$.

@@

	While the discrepancy between the experimental result and the analytical model result grew as the height increased the difference stayed negligible but the difference compared to the vacuum model grew. At $8.68\,m$ the vacuum model is no longer a viable option to accurately describe the path of the object. The difference $0.039\,s$ can be partially accounted for by the difficulty in determining the exact time at which the object was dropped. Instead of using a human to drop the object and a human to film the sensor a better alternative would be a electronically controlled holder connected to a computer and a photosensor at the bottom, also connected to the same computer. The coefficient of drag of a sphere is also hard to determine exactly. Some\cite{c_d} say the coefficient is $0.5$ instead of $0.47$ which decreases the discrepancy for the second fall by roughly $25\,\%$. Redoing the experiment in a lab could improve the accuracy as minor winds or unpredictable levels or air density could be avoided. In conclusion it is impossible to say the analytical air resistance model developed has a fundamental error and if any further tweaking were to take place a deeper analysis of the thrown object would be required. The numerical model approached the analytical one in terms of results as $\Delta \rightarrow 0$.

@@

>>>>

	The time to highest point is the amount of time it takes for our object to reach its peak height which we will denote $t_{peak}$. Since $y(t)$ is a second degree polynominal it only has one extremum. This must be the maximum because $v_y(0) > 0$. We can therefore solve $v_y = 0$ for $t_{peak}$.

@@

	We can plot another $y(t)$ where $k$ is much smaller which should result in our object not reaching its terminal velocity in such short time. The following measurements are from a small steel ball with $k = 0.00313$.
	\begin{align}\nonumber
		\begin{cases}\nonumber
			C_D &= 0.47 \\
			m &= 0.022\,kg\\
			y_0 &= 3\,m\\
			v_0 &= 0\,m/s\\
			\rho &= 1.225\,kg/m^3\\
			A &= 0.00023916\,m^2
		\end{cases}
	\end{align}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ xlabel=t (s), ylabel=y (m)]
		\addplot[solid] coordinates {(0.000, 3.000) (0.010, 3.000) (0.020, 2.998) (0.030, 2.996) (0.040, 2.992) (0.050, 2.988) (0.060, 2.982) (0.070, 2.976) (0.080, 2.969) (0.090, 2.960) (0.100, 2.951) (0.110, 2.941) (0.120, 2.929) (0.130, 2.917) (0.140, 2.904) (0.150, 2.890) (0.160, 2.874) (0.170, 2.858) (0.180, 2.841) (0.190, 2.823) (0.200, 2.804) (0.210, 2.784) (0.220, 2.762) (0.230, 2.740) (0.240, 2.717) (0.250, 2.693) (0.260, 2.668) (0.270, 2.642) (0.280, 2.615) (0.290, 2.587) (0.300, 2.558) (0.310, 2.528) (0.320, 2.497) (0.330, 2.466) (0.340, 2.433) (0.350, 2.399) (0.360, 2.364) (0.370, 2.328) (0.380, 2.292) (0.390, 2.254) (0.400, 2.215) (0.410, 2.175) (0.420, 2.135) (0.430, 2.093) (0.440, 2.050) (0.450, 2.007) (0.460, 1.962) (0.470, 1.917) (0.480, 1.870) (0.490, 1.823) (0.500, 1.774) (0.510, 1.725) (0.520, 1.674) (0.530, 1.623) (0.540, 1.570) (0.550, 1.517) (0.560, 1.463) (0.570, 1.407) (0.580, 1.351) (0.590, 1.294) (0.600, 1.236) (0.610, 1.176) (0.620, 1.116) (0.630, 1.055) (0.640, 0.993) (0.650, 0.930) (0.660, 0.866) (0.670, 0.801) (0.680, 0.735) (0.690, 0.668) (0.700, 0.600) (0.710, 0.531) (0.720, 0.461) (0.730, 0.391) (0.740, 0.319) (0.750, 0.246) (0.760, 0.172) (0.770, 0.098) (0.780, 0.022) (0.790, -0.055) };

@@

	In contrary to our LAT approximation in section \ref{seq:lat} we must now separate the cases where $\hat{w} = \hat{\jmath}, \hat{w} = -\hat{\jmath}$ because a negative $w$ should not lead to a positive $u'$. When $v$ was approximated as $u$ we assumed that $\hat{u} = \hat{\imath}$ because it is the only realistic scenario. Now that $v$ is approximated as $w$ we need to consider its direction before that context. In the end our equations of motions will consist of two separate $u$ and $w$.

@@

	While the discrepancy between the experimental result and the analytical model result grew as the height increased the difference stayed negligible but the difference compared to the vacuum model grew. At $8.68\,m$ the vacuum model is no longer a viable option to accurately describe the path of the object. The difference $0.04\,s$ can be partially accounted for by the difficulty in determining the exact time at which the object was dropped. Instead of using a human to drop the object and a human to film the sensor a better alternative would be a electronically controlled holder connected to a computer and a photosensor at the bottom, also connected to the same computer. The coefficient of drag of a sphere is also hard to determine exactly. Some\cite{c_d} say the coefficient is $0.5$ instead of $0.47$ which decreases the discrepancy for the second fall by roughly $25\,\%$. Redoing the experiment in a lab could improve the accuracy as minor winds or unpredictable levels or air density could be avoided. In conclusion it is impossible to say the analytical air resistance model developed has a fundamental error and if any further tweaking were to take place a deeper analysis of the thrown object would be required. The numerical model approached the analytical one in terms of results as $\Delta \rightarrow 0$.

@@

		\bibitem{riseAndFall} 
		Peter Timmerman, Jacobus P. van der Weele \textit{On the rise and fall of a ball with linear or quadratic drag}.
		Centre for Theoretical Physics, University of Twente
		Am.J.Phys. 67, 538 (1999)
		\\\url{http://dx.doi.org/10.1119/1.19320}
